# Results

# Brooke's results

## Hypothesis

As a reminder, this project hypothesized that both the intensity and involvement factors will predict song popularity such that the higher the factor score, the more likely the the song will have a higher popularity score (over and above the effects of the other factor scores). The third factor, technicality, will not significantly predict song popularity.

```{r, include = F}
# loading packages
library(tidyverse)
library(psych)
getwd()
# loading data
d <- read.csv("/Users/brookesasia/Desktop/UW-Madison/PHS 650/sarah_brooke_final_paper/Homework3_variables.csv")

spotify_songs <- tidytuesdayR::tt_load('2020-01-21')$spotify_songs

# checking variable names
names(d)
names(spotify_songs)

# keeping track_popularity and track_id to merge into d
spotify_songs <- spotify_songs %>%
  select(track_id, track_popularity)

# merging d and spotify_songs
d <- d %>%
  full_join(spotify_songs, by = "track_id")

names(d)
```

## Linear model

```{r}
m_pop <- lm(track_popularity ~ pa1_intensity + pa2_involvement + pa3_technicality, data = d)

m_pop_results <- summary(m_pop)

m_pop_results

# notes for later
# save the summary into a variable, X. then you have X as a data frame and you can treat it like a dataframe where you create variables and lower confidence interval is b minus SE and create it as a variable and then do the same thing for the upper. then you could also round all of the variables and select out what you do and don't want. and then write it kable whatever when you have the way you want it to look.
```
